---
title: "ACTUR-SEER Data Consolidation"
author: "Darryl Nousome"
date: "12/31/2020"
output: 
  word_document:
    reference_docx: "/home/dnousome/Documents/margins.docx"

---

```{r setup, include=FALSE,echo=F}
knitr::opts_chunk$set(echo = TRUE)
library(readxl)
library(kableExtra)
library(tidyverse)
library(arsenal)
library(haven)
dt_final=read_rds("~/Documents/G/Work_Bench/Epidemiology/Projects/MCC_CPDR SEERACTUR 2020/Data/mcc_actur.rds")
labelsdt=list(year_cat="Year of Diagnosis",age="Age at diagonsis",age_cat5="Age Categories",race_new="Race",
              stage_final1="Tumor Stage",tumor_grade_gleason="Tumor Grade Consolidated",
              PSA="PSA",PSA_category="PSA Positive/Negative",PSA_cat="PSA Categories",
              radiation_new="Consolidated Radiation",
              CSsitespecificfactor12004varyi="SEER PSA",Gleason_prostatectomy="Gleason from Prostatectomy",
              Gleason_biopsy="Gleason from Biopsy", Gleason_any="Gleason received",Gleason_final="Final Gleason Score",Gleason_final_method="Final Gleason Score from Method")

```

## Variable Cleaning
Matching variables (age, race, diagnosis year) were included in the dataset. Additional variables including surgery, radiation, PSA, Gleason score will be processed and consolidated.  

## Surgery
```{r surg,echo=F,results="asis"}
d=tableby(dataset~RX_Summ__Surg_Prim_Site+Sitespecificsurgery19731997var+RX_Summ_Surg_Prim_Site_1998,data=dt_final,total=F,test=F) 
summary(d,title="Original Surgery Variables from ACTUR and SEER") 
```


For both ACTUR and SEER  
00 replaced to No surgery  
99/Blank -> Unknown/Missing  
14-90-> Yes
  
For SEER, the variable for surgery from 1973-1997 (Site specific surgery (1973-1997 varying)):  
01, 02, 03, 05, 06, 07-> No surgery  

Referenced https://seer.cancer.gov/archive/manuals/historic/AppendD.pdf
  
  
For SEER 1998+, (RX_Summ_Surg_Prim_Site_1998 variable) used the NAACR and SEER 2003 guidelines

For Reference:  
Original NAACR  
00: None  
10-19: Site-specific codes; tumor destruction  
20-80: Site-specific codes; resection  
90: Surgery, NOS  
98: Site-specific codes; special  
99: Unknown  

## Radiation
```{r radiation,echo=F,results="asis"}
d=tableby(dataset~Radiation+RX_Summ__Radiation,data=dt_final,total=F,test=F) 
summary(d,title="Original Radiation Variables from ACTUR and SEER") 
```
SEER: Beam radiation, Combination of beam with implants or isotopes, Radiation, NOS  method or source not specified, Radioactive implants, Radioisotopes- Coded to Yes  
None/refused-> No  
Unknown, Recommended, unknown if administered->  Unknown/Missing  
  
ACTUR: 1-6 Coded to Yes  
0,7-> No  
8,9 or Blank->Unknown/Missing

Original NAACR  
0: None  
1: Beam radiation  
2: Radioactive implants  
3: Radioisotopes  
4: Combination of 1 with 2 or 3  
5: Radiation, NOS—method or source not specified  
7: Patient or patient's guardian refused*  
8: Radiation recommended, unknown if administered*  
9: Unknown if radiation therapy administered  

## Gleason  
```{r gleason,echo=F,results="asis"}
dt_g=dt_final %>% mutate(Gleason_biopsy=ifelse(is.na(Gleason_biopsy),"NA",Gleason_biopsy)) %>%
  mutate(Gleason_prostatectomy=ifelse(is.na(Gleason_prostatectomy),"NA",Gleason_prostatectomy)) %>%
  mutate(Gleason_biopsy=factor(Gleason_biopsy,levels=c("NA","6","7","8","9","10"))) %>%
  mutate(Gleason_prostatectomy=factor(Gleason_prostatectomy,levels=c("NA","6","7","8","9","10")))


d=tableby(Gleason_biopsy~Gleason_prostatectomy,data=dt_g,total=F,test=F) 

summary(d,title="Gleason Biopsy/Prostatectomy") 


#library(xtable)
#df <- data.frame(vertical_title = dt_g$Gleason_biopsy, horitzontal_title = dt_g$Gleason_prostatectomy)
#df <- as.data.frame.matrix(table(df)) 


#print(xtable(df,
          #  align = "|l|rrr|",
#            caption = "Test"),
#      caption.placement = 'top',
#      comment = F,
#      scalebox=1,
#      include.rownames = T,
#      hline.after = c(-1,0,nrow(df)),
#      vline.after = c(1),
#      format.args=list(big.mark = ",", decimal.mark = "."))

```  
A large majority are missing Gleason scores, however, more are missing in the prostatectomy group.


ACTUR/SEER: 988,998,999 Coded to Missing  
000, 002, 003,004,005 -> Missing since these are not valid scores  
A new consolidated variable for Gleason was created taking the prostatectomy score first followed by biopsy

Examine the Gleason score by Year
```{r gleason_time,echo=F,results="asis"}


d=tableby(year_cat~Gleason_any+Gleason_prostatectomy+Gleason_biopsy+Gleason_final+Gleason_final_method,data=dt_final,total=F,test=T) 
summary(d,title="Gleason Biopsy/Prostatectomy split by time",labelTranslations=labelsdt) 

dt_a=dt_final %>% filter(year_cat %in% c("2000-2004","2005-2009","2010-2013") & dataset=="ACTUR")
d=tableby(year_cat~Gleason_any+Gleason_prostatectomy+Gleason_biopsy+Gleason_final+Gleason_final_method,data=dt_a,total=F,test=T) 
summary(d,title="Gleason Biopsy/Prostatectomy in ACTUR",labelTranslations=labelsdt) 

dt_s=dt_final %>% filter(year_cat %in% c("2000-2004","2005-2009","2010-2013") & dataset=="SEER")
d=tableby(year_cat~Gleason_any+Gleason_prostatectomy+Gleason_biopsy+Gleason_final+Gleason_final_method,data=dt_s,total=F,test=T) 
summary(d,title="Gleason Biopsy/Prostatectomy in SEER",labelTranslations=labelsdt) 
```  

### Compare Gleason 
Compare Gleasons across methods for patients that have both Gleason types
n=1501 for participants that have both.
```{r gleacomp,echo=F,results='asis'}
t1=dt_final %>% filter(!is.na(Gleason_biopsy),!is.na(Gleason_prostatectomy)) %>% select(Gleason_biopsy,Gleason_prostatectomy,pair,dataset) %>%
  mutate(ID=paste0(pair,"_",dataset)) %>% mutate(ID=make.unique(ID)) 
t2=pivot_longer(t1,c(Gleason_biopsy,Gleason_prostatectomy))
ggplot(t2,aes(x=name,fill=factor(value,levels=c(10:6)))) +geom_bar(position="fill") + labs(fill = "Gleason Score") +xlab("Gleason Type")
tt=t.test(t1$Gleason_biopsy,t1$Gleason_prostatectomy,paired=T)
d=tableby(name~value,data=t2,total=F,test=T) 
summary(d,title="Gleason in those with both Biopsy and Prostatectomy") 
```

### Final Tumor Grade Consolidated Gleason 
Additionally for Gleason scores, we will use the Gleason scores first to identify the Tumor Grade (Pathologic examination determines the grade https://seer.cancer.gov/tools/grade/).
If we are unable to set the tumor grade through Gleason from 2010-2013, then use the original grade given from the database.  

```{r gleagrade,echo=F,results='asis'}

d=tableby(dataset~tumor_grade_gleason,data=dt_final,total=F,test=T) 
summary(d,title="Final Grade/Gleason",labelTranslations=labelsdt) 


```


### PSA
Similar approach with PSA
PSA split by time  
```{r psa,echo=F,results="asis"}
d=tableby(year_cat~PSA+PSA_category,data=dt_final,total=F,test=T) 
```

```{r psaactur,echo=F,results="asis"}
d=tableby(year_cat~PSA+PSA_category,data=dt_a,total=F,test=T) 
summary(d,title="PSA in ACTUR") 
```

```{r psaseer,echo=F,results="asis"}
d=tableby(year_cat~PSA+PSA_category,data=dt_s,total=F,test=T) 
summary(d,title="PSA in SEER") 
```  

More missing PSA in the SEER Group, notably during 2000-2009. To be comparable, PSA from ACTUR will not be included as well until 2010.