---
title: "ACTUR-SEER Data Analysis"
author: "Darryl Nousome"
date: "12/31/2020"
output: #html_document
  word_document:
    reference_docx: "/home/dnousome/Documents/margins.docx"

---

```{r setup, include=FALSE,echo=F}
knitr::opts_chunk$set(echo = TRUE)
library(readxl)
library(kableExtra)
library(tidyverse)
library(arsenal)
library(haven)


labelsdt=list(year_cat="Year of Diagnosis",age="Age at diagonsis",age_cat5="Age Categories",race_new="Race",
              stage_final1="Tumor Stage",tumor_grade_gleason="Tumor Grade",age_cat1="Age Categories",
              PSA="PSA",PSA_category="PSA Positive/Negative",PSA_cat="PSA Categories",PSA_cat_rem="PSA Categories",
              radiation_new="Consolidated Radiation",
              CSsitespecificfactor12004varyi="SEER PSA",Gleason_prostatectomy="Gleason from Prostatectomy",
              Gleason_biopsy="Gleason from Biopsy", Gleason_any="Gleason received",
              Gleason_final="Final Gleason Score",Gleason_final_rem="Final Gleason Score",
              Gleason_final_method1="Final Gleason Score from Method")

dt_final=read_rds("~/Documents/G/Work_Bench/Epidemiology/Projects/MCC_CPDR SEERACTUR 2020/Data/mcc_actur.rds") %>%
  mutate(dataset=factor(dataset,levels=c("SEER","ACTUR"))) %>%
  mutate(Dxyr_PSA_all=factor(Dxyr_PSA,levels=c("0","1","2","3","8","9"),
                         labels=c("PSA 0-<4","PSA 4-<10","PSA 10-<20","PSA \u226520",
                                  "PSA Missing in 2010-2013","PSA Missing any other time"))) %>%
  mutate(PSA_cat_rem=factor(PSA_cat_rem,levels = c("1","2","3","4"),labels=c("0-<4","4-<10","10-<20","\u226520"))) %>%
  mutate(Gleason_final_rem=cut(Gleason_final_rem,breaks =c(6,7,8,11),labels=c("6","7","8-10"),right=F)) %>%
  mutate(Gleason_final_method1=ifelse(is.na(Gleason_final_rem),NA,Gleason_final_method))
 

```


## Summary tables
```{r tables,echo=F}
tab.noby <- tableby(dataset~ age+age_cat1+year_cat+race_new+Hispanic+stage_final1+tumor_grade_gleason+PSA+PSA_category+PSA_cat_rem+
                      Gleason_final_rem+Gleason_final_method1+Surgery+radiation_new, data=dt_final,test=T,total=F) 

summary(tab.noby,labelTranslations=labelsdt,digits=1) %>% 
  kable(format="markdown") %>%
  add_footnote(c("PSA values and Gleason scores for SEER were available from 2010-2013",
                 "Gleason score was used to determine Tumor Grade when available"),notation = "none")

```

## Overall death and five-year survival
162 deaths occur after 2013 and are coded as non-deaths in ACTUR in the time period for overall survival  
176 with deaths however, they do not have death dates in DEERS their date date will be recoded to date of last contact  
2 with deaths before dx date in ACTUR: IDs 3703, 5212. Respective pairs in SEER will be dropped as well  

Five year survival was calculated and log-rank survival curves were tested
```{r surv,echo=F,warning=F,message=F, fig.width=10,fig.height=8}
dt_surv=read_rds("../Data/mcc_actur_surv.rds") %>%

  mutate(Dxyr_PSA_all=factor(Dxyr_PSA_all,levels=c("0","1","2","3","8","9"),
                            labels=c("PSA 0-<4","PSA 4-<10","PSA 10-<20","PSA \u226520",
                                  "PSA Missing in 2010-2013","PSA Missing any other time"))) %>%
  mutate(PSA_cat_rem=factor(PSA_cat_rem,levels = c("1","2","3","4"),labels=c("0-<4","4-<10","10-<20","\u226520"))) %>%
  mutate(dataset=factor(dataset,levels=c("SEER","ACTUR")))

library(survival)
library(survminer)
fit <- survfit(Surv(fiveyear_FU, fiveyear_death) ~ dataset,
               data = dt_surv)
ggsurvplot(fit, data = dt_surv, risk.table = 'nrisk_cumevents', conf.int = T,censor=F,pval=T,fontsize=4,tables.y.text=F, legend.labs = c("SEER", "ACTUR"),xlab="Time (Months)",
           title="Five Year Survival Between ACTUR compared to SEER")


dt_surv_t3=dt_surv %>% filter(stage_final==3)
fit <- survfit(Surv(fiveyear_FU, fiveyear_death) ~ dataset,
               data = dt_surv_t3)
ggsurvplot(fit, data = dt_surv_t3, risk.table = 'nrisk_cumevents', conf.int = T,censor=F,pval=T,fontsize=4,tables.y.text=F, legend.labs = c("SEER","ACTUR"),xlab="Time (Months)",
                    title="Five Year Survival Between ACTUR compared to SEER (T3 Stage only)")

dt_surv_t4=dt_surv %>% filter(stage_final==4)
fit <- survfit(Surv(fiveyear_FU, fiveyear_death) ~ dataset,
               data = dt_surv_t4)
ggsurvplot(fit, data = dt_surv_t4, risk.table = 'nrisk_cumevents', conf.int = T,censor=F,pval=T,fontsize=4,tables.y.text=F, legend.labs = c("SEER", "ACTUR"),xlab="Time (Months)",
                    title="Five Year Survival Between the ACTUR compared to SEER (T4 Stage only)")
```



### Cox Proportional Hazard Modeling
Used a multivariable Cox proportional hazards model to estimate hazard ratios (HRs) and 95% confidence intervals (CIs) for 5-year survival between ACTUR cases and SEER cases. Stratified on Matching ID to account for matched data. Schoenfeld residuals were examined to fufill assumptions of PH.
As mentioned in the data consolidation results, Gleason and PSA were used only from 2010 onward for both datasets. For Gleason, the prostatectomy score was preferred over the biopsy score. An indicator was created for those years prior to 2010 to ensure sample size was preserved.   
```{r cox,echo=F,message=F,warning=F}


#dt_surv_check=dt_surv %>% filter(DEERS_Death_Date=="" & death_new==1 &dataset=="ACTUR")


labels_coxph=c(labelsdt,
           `datasetACTUR`="ACTUR vs SEER",
           `stage_final14`="Stage 4 vs Stage 3",

           `SurgeryUnknown/Missing`="Surgery Unknown vs No",
           `SurgeryYes`="Surgery Yes vs No",
           `radiation_newUnknown/Missing`="Radiation Unknown vs No",
           `radiation_newYes`="Radiation Yes vs No",
           `age`="Age (Continuous)",
           `as.factor(hispanic)1`="Hispanic vs Non-Hispanic",
           `as.factor(hispanic)9`="Hispanic vs Unknown Hispanic",
           `PSA_cat[4,10)`="PSA 4-<10 vs <4",
           `PSA_cat[10,20)`="PSA 10-<20 vs <4",
           `PSA_cat[20,96)`="PSA \u226520 vs <4",
          `as.factor(Dxyr_PSA_all)PSA 4-<10`="PSA 4-<10 vs 0-<4",
          `as.factor(Dxyr_PSA_all)PSA 10-<20`="PSA 10-<20 vs 0-<4",
          `as.factor(Dxyr_PSA_all)PSA ≥20`="PSA \u226520 vs 0-<4",
          `as.factor(Dxyr_PSA)5`="PSA Missing in 2010-2013 vs PSA 0-<4",
          `as.factor(Dxyr_PSA)6`="All Other years PSA Missing vs PSA 0-<4",
`as.factor(Dxyr_PSA_all)PSA Missing in 2010-2013`="PSA Missing in 2010-2013 vs PSA 0-<4",
`as.factor(Dxyr_PSA_all)PSA Missing any other time`="All Other years PSA Missing vs PSA 0-<4",
           `tumor_grade_gleason2`="Grade 2 vs Grade 1",
           `tumor_grade_gleason3`="Grade 3 vs Grade 1",
          `tumor_grade_gleason4`="Grade 4 vs Grade 1",
            `tumor_grade1Unknown/Missing`="Unknown vs Grade 1",
            `tumor_grade12`="Grade 2 vs Grade 1",
            `tumor_grade13`="Grade 3 vs Grade 1",
            `tumor_grade14`="Grade 4 vs Grade 1",
            `tumor_grade_gleasonUnknown/Missing`="Unknown vs Grade 1",
          
            `tumor_grade_gleasonMissing/Unknown`="Missing Grade vs 1",
           Gleason_ind="Gleason indicator variable",
           PSA_cat_ind="PSA indicator variable")

cont=modelsum.control(survival.stats=c("HR", "CI.lower.HR", "CI.upper.HR", "p.value"))

summary(modelsum(Surv(fiveyear_FU, fiveyear_death) ~dataset,adjust=~stage_final1+tumor_grade_gleason+as.factor(Dxyr_PSA_all)+
                   Surgery+radiation_new+age+as.factor(hispanic)+strata(pair),
                 data=dt_surv,family="survival",control=cont),
       labelTranslations=labels_coxph,digits=2,digits.p=3,digits.ratio=2) %>%
  kable(format="markdown") %>%
  add_footnote(c("Missing Tumor Grade was replaced by Gleason score when available"),notation = "none")



fit1=coxph(Surv(fiveyear_FU, fiveyear_death) ~dataset+stage_final1+tumor_grade_gleason+as.factor(as.numeric(Dxyr_PSA_all))+
                   Surgery+radiation_new+age+as.factor(hispanic)+
                 
                    strata(pair),data=dt_surv)
ggcoxdiagnostics(fit1, type = "schoenfeld")
  
```

### Stratified Analyses  
Stratified analyses were performed by restricting to variables that were significantly associated with survival times to examine if there were differences in that variable between ACTUR and SEER. 
Models were adjusted for Grade, Stage, Surgery, Radiation, Age, Hispanic ethnicity, PSA, Gleason, and matched. The variable was not included when the model was stratified.
```{r strat,echo=F,message=F,warning=F}

##STRAT
surv1=filter(dt_surv,stage_final1==3)
summary(modelsum(Surv(fiveyear_FU, fiveyear_death) ~dataset,adjust=~tumor_grade_gleason+as.factor(as.numeric(Dxyr_PSA_all))+
                   Surgery+radiation_new+age+as.factor(hispanic)+strata(pair),
                 data=surv1,family="survival",show.adjust=F,control=cont),
       labelTranslations=labels_coxph,digits=2,digits.p=3,digits.ratio=2) %>%
  kable(format="markdown",caption="Stage 3")

surv1=filter(dt_surv,stage_final1==4)
summary(modelsum(Surv(fiveyear_FU, fiveyear_death) ~dataset,adjust=~tumor_grade_gleason+as.factor(as.numeric(Dxyr_PSA_all))+
                   Surgery+radiation_new+age+as.factor(hispanic)+strata(pair),
                 data=surv1,family="survival",show.adjust=F,control=cont),
       labelTranslations=labels_coxph,digits=2,digits.p=3,digits.ratio=2) %>%
  kable(format="markdown",caption="Stage 4")

##Grade
#surv1=filter(dt_surv,tumor_grade1==1)
#summary(modelsum(Surv(fiveyear_FU, fiveyear_death) ~dataset,adjust=~stage_final1+Surgery+radiation_new+age+as.factor(hispanic)+strata(pair)+as.factor(PSA_cat)+as.factor(Gleason_final),
#                 data=surv1,family="survival",show.adjust=F),
#       labelTranslations=labels_coxph,digits=2,digits.p=3,digits.ratio=2) %>%
#  kable(format="markdown",caption="Grade 1")

surv1=filter(dt_surv,tumor_grade_gleason==1)
summary(modelsum(Surv(fiveyear_FU, fiveyear_death) ~dataset,adjust=~stage_final1+as.factor(as.numeric(Dxyr_PSA_all))+
                   Surgery+radiation_new+age+strata(pair),
                 data=surv1,family="survival",show.adjust=F,control=cont),
       labelTranslations=labels_coxph,digits=2,digits.p=3,digits.ratio=2) %>%
  kable(format="markdown",caption="Grade 1")

surv1=filter(dt_surv,tumor_grade_gleason==2)
summary(modelsum(Surv(fiveyear_FU, fiveyear_death) ~dataset,adjust=~stage_final1+as.factor(as.numeric(Dxyr_PSA_all))+
                   Surgery+radiation_new+age+as.factor(hispanic)+strata(pair),
                 data=surv1,family="survival",show.adjust=F,control=cont),
       labelTranslations=labels_coxph,digits=2,digits.p=3,digits.ratio=2) %>%
  kable(format="markdown",caption="Grade 2")

surv1=filter(dt_surv,tumor_grade_gleason==3)
summary(modelsum(Surv(fiveyear_FU, fiveyear_death) ~dataset,adjust=~stage_final1+as.factor(as.numeric(Dxyr_PSA_all))+
                   Surgery+radiation_new+age+as.factor(hispanic)+strata(pair),
                 data=surv1,family="survival",show.adjust=F,control=cont),
       labelTranslations=labels_coxph,digits=2,digits.p=3,digits.ratio=2) %>%
  kable(format="markdown",caption="Grade 3")


surv1=filter(dt_surv,tumor_grade_gleason==4)
summary(modelsum(Surv(fiveyear_FU, fiveyear_death) ~dataset,adjust=~as.factor(as.numeric(Dxyr_PSA_all))+
                   Surgery+age+strata(pair),
                 data=surv1,family="survival",show.adjust=F,control=cont),
       labelTranslations=labels_coxph,digits=2,digits.p=3,digits.ratio=2) %>%
  kable(format="markdown",caption="Grade 4")


##Age
surv1=filter(dt_surv,age_cat1=="<50")
summary(modelsum(Surv(fiveyear_FU, fiveyear_death) ~dataset,adjust=~stage_final1+tumor_grade_gleason+as.factor(as.numeric(Dxyr_PSA_all))+
                   Surgery+as.factor(hispanic)+strata(pair),
                 data=surv1,family="survival",show.adjust=F,control = cont),
       labelTranslations=labels_coxph,digits=2,digits.p=3,digits.ratio=2) %>%
  kable(format="markdown",caption="Age <50")



surv1=filter(dt_surv,age_cat1=="50-<65")
summary(modelsum(Surv(fiveyear_FU, fiveyear_death) ~dataset,adjust=~stage_final1+tumor_grade_gleason+as.factor(as.numeric(Dxyr_PSA_all))+
                   Surgery+as.factor(hispanic)+strata(pair),
                 data=surv1,family="survival",show.adjust=F,control=cont),
       labelTranslations=labels_coxph,digits=2,digits.p=3,digits.ratio=2) %>%
  kable(format="markdown",caption="Age 50-<65")

surv1=filter(dt_surv,age_cat1=="≥65")
summary(modelsum(Surv(fiveyear_FU, fiveyear_death) ~dataset,adjust=~stage_final1+tumor_grade_gleason+as.factor(as.numeric(Dxyr_PSA_all))+
                   Surgery+as.factor(hispanic)+strata(pair),
                 data=surv1,family="survival",show.adjust=F,control=cont),
       labelTranslations=labels_coxph,digits=2,digits.p=3,digits.ratio=2) %>%
  kable(format="markdown",caption="Age ≥65")


##Race
surv1=filter(dt_surv,race_new=="Asian or Pacific Islander")
summary(modelsum(Surv(fiveyear_FU, fiveyear_death) ~dataset,adjust=~stage_final1+tumor_grade_gleason+as.factor(as.numeric(Dxyr_PSA_all))+
                   Surgery+as.factor(hispanic)+strata(pair),
                 data=surv1,family="survival",show.adjust=F,control=cont),
       labelTranslations=labels_coxph,digits=2,digits.p=3,digits.ratio=2) %>%
  kable(format="markdown",caption="Race Asian or Pacific Islander")

 surv1=filter(dt_surv,race_new=="Black")
summary(modelsum(Surv(fiveyear_FU, fiveyear_death) ~dataset,adjust=~stage_final1+tumor_grade_gleason+as.factor(as.numeric(Dxyr_PSA_all))+
                   Surgery+as.factor(hispanic)+strata(pair),
                 data=surv1,family="survival",show.adjust=F,control=cont),
       labelTranslations=labels_coxph,digits=2,digits.p=3,digits.ratio=2) %>%
  kable(format="markdown",caption="Race Black")


surv1=filter(dt_surv,race_new=="White")
summary(modelsum(Surv(fiveyear_FU, fiveyear_death) ~dataset,adjust=~stage_final1+tumor_grade_gleason+as.factor(as.numeric(Dxyr_PSA_all))+
                   Surgery+as.factor(hispanic)+strata(pair),
                 data=surv1,family="survival",show.adjust=F,control=cont),
       labelTranslations=labels_coxph,digits=2,digits.p=3,digits.ratio=2) %>%
  kable(format="markdown",caption="White")



                                           


##SURG
surv1=filter(dt_surv,Surgery=="Yes")
summary(modelsum(Surv(fiveyear_FU, fiveyear_death) ~dataset,adjust=~stage_final1+tumor_grade_gleason+as.factor(as.numeric(Dxyr_PSA_all))+
                   radiation_new+age+as.factor(hispanic)+strata(pair),
                 data=surv1,family="survival",show.adjust=F,control=cont),
       labelTranslations=labels_coxph,digits=2,digits.p=3,digits.ratio=2) %>%
  kable(format="markdown",caption="Surgery Yes")

surv1=filter(dt_surv,Surgery=="No")
summary(modelsum(Surv(fiveyear_FU, fiveyear_death) ~dataset,adjust=~stage_final1+tumor_grade_gleason+as.factor(as.numeric(Dxyr_PSA_all))+
                   radiation_new+age+as.factor(hispanic)+strata(pair),
                 data=surv1,family="survival",show.adjust=F,control=cont),
       labelTranslations=labels_coxph,digits=2,digits.p=3,digits.ratio=2) %>%
  kable(format="markdown",caption="Surgery No")



##Radiation
surv1=filter(dt_surv,radiation_new=="Yes")
summary(modelsum(Surv(fiveyear_FU, fiveyear_death) ~dataset,adjust=~stage_final1+tumor_grade_gleason+as.factor(as.numeric(Dxyr_PSA_all))+
                   Surgery+age+as.factor(hispanic)+strata(pair),
                 data=surv1,family="survival",show.adjust=F,control=cont),
       labelTranslations=labels_coxph,digits=2,digits.p=3,digits.ratio=2) %>%
  kable(format="markdown",caption="Radiation Yes")

surv1=filter(dt_surv,radiation_new=="No")
summary(modelsum(Surv(fiveyear_FU, fiveyear_death) ~dataset,adjust=~stage_final1+tumor_grade_gleason+as.factor(as.numeric(Dxyr_PSA_all))+
                   Surgery+age+as.factor(hispanic)+strata(pair),
                 data=surv1,family="survival",show.adjust=F,control=cont),
       labelTranslations=labels_coxph,digits=2,digits.p=3,digits.ratio=2) %>%
  kable(format="markdown",caption="Radiation No")

```

